---
title: "World Bank List of Economies and Groupings"
author: "Jane Pratt"
date: "07/04/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    theme: cosmo
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this document we'll retrieve the World Bank's list of economies broken down by region and income, as well as other groupings used by the World Bank. 

## Goal

The goal here is simply to obtain data that will be useful for other projects.

## Libraries

```{r, message = FALSE}
library(tidyverse)
```

## Data Source

Information about the data and links to download the files are available at:
https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups

Additional data sets are available such as historic data about the income groups for each economy.


### License Information

No specific license is mentioned on the download page for this data, so the default license for World Bank Data applies:

[Creative Commons Attribution 4.0 (CC-BY 4.0)](https://creativecommons.org/licenses/by/4.0/)


### Notes

The data includes the following:

- Economy - name of the economy (which is not necessarily  a sovereign nation - eg it could be a dependent territory)
- Country Code - 3 character code used to refer to this economy by the World Bank. ISO 3166-1 codes are used where available
- Region - name of the region in which the economy is located
- Income Group - category based on the Gross National Income per capita (income levels for each group are specified on the download page)
- Lending Category - eligibility for loans and support from the IDA (International Development Association), IBRD (International Bank for Reconstruction and Development) or Blend (a combination of the two)
- Other:
  * HIPC - Heavily Indebted Poor Countries
  * EMU - Member of EU Economic & Monetary Union
- Groups - a number of other country groupings based on region, demographics, income etc
  * Group code
  * Group Name
  * Country code
  * Country name

# Loading the data

Let's start by downloading the latest data:

```{r, message = FALSE, warning = FALSE}
economies_source_xl <- 'http://databank.worldbank.org/data/download/site-content/CLASS.xls'
economies_local_xl <- 'Data/wb_economies_class.xls'
download.file(economies_source_xl, economies_local_xl, mode = 'wb', quiet = TRUE)
```

## List of economies

We can read in the data using the readxl package from the tidyverse. We'll need to skip the first 4 rows on the 'List of economies' sheet as there are some titles above the main table.

```{r, message = FALSE}
economies_data_all <- readxl::read_excel(economies_local_xl, sheet = 'List of economies', skip = 4)
economies_data_all
```

The 'List of economies' sheet contains a table with the individual economies at the top, but below this is a section with the name and code for the groupings, then a list of notes about the data. Only the individual economies contain data in the 'Region' column.

There are also blank columns in the worksheet. 

Let's remove the unwanted rows and columns:

```{r}
economies_data <- economies_data_all %>%
  select(-c(x...1, x...2, X)) %>% # Remove empty columns
  filter(!Economy == 'x') %>% # Remove row of x's at the top of the table
  filter(!is.na(Region)) %>% # Remove extra data from below the main table which have no region
  mutate(Region = as.factor(Region)) %>%
  mutate('Income group' = as.factor(`Income group`))
economies_data
write.csv(economies_data, 'Data/wb_economies.csv')
```

```{r}
wb_economy_groupings <- readxl::read_excel(economies_local_xl, sheet = 'Groups')
write.csv(wb_economy_groupings, 'Data/wb_groups.csv')
wb_economy_groupings
```

# Further Information

- [World Bank Country and Lending Groups](https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)
- [World Bank Data Access and Licensing](https://datacatalog.worldbank.org/public-licenses)


